---
title: "Diagnosticity of p values in the face of ANOVA's hidden multiplicity"
author: "Ian Hussey^[Ghent University. Email: ian.hussey@ugent.be]"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r, include=FALSE}

# set defaults
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE)

```

```{r}

# dependencies
library(tidyverse)

# disable scientific notation
options(scipen = 999) 

# knitr output for html (comment out if knitting to pdf)
options(knitr.table.format = "html")

```

# True false positive rate (FPR)

Imagine we're running a 7X2X2X2X2 between participants experiment. This is a real world example.

Setting alpha to its traditional 0.05 might imply that we expect or will tolerate a false positive rate of 0.05.

However, given that ANOVAs involve the testing of multiple main and interaction effects, and results are rarely subjected to multiple testing corrections, the FPR is typically much higher than this. But by how much?

A 7X2X2X2X2 deign involves:

- 5 main effects
- 10 two-way interaction effects
- 10 three-way interaction effects
- 5 four-way interaction effects
- 1 five-way interaction effect

= 31 total tests

This can be simplified as n_tests = 2^factors-1, where design is the number of factors in the design (i.e., 5 here).

Given an alpha value of 0.05, the FPR for between groups designs can be shown mathmathically to be (1-alpha)^n_tests =  (Craemers et al., 2015).

```{r}

fpr_given_hidden_multiplicity <- function(alpha, factors){
  (1 - alpha)^(2^(factors)-1)
}

fpr_given_hidden_multiplicity(0.05, 5)

```
